Chasing \rainbows and ocean glints:\\ Inner working angle constraints for the Habitable Worlds Observatory

### 2023-07-27 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{abstract}
NASA is engaged in planning for a Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones and establish their habitability. 
%
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
%
Direct imaging missions are optimised for planets near quadrature, but \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
%
The range of accessible phase angles for a given exoplanet will depend on the planet's orbital inclination and/or the coronagraph's inner working angle (IWA). 
%
We use a recently-created catalog relevant to HabWorlds of 164 stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and polarization effects due to Rayleigh scattering. 
%
We find that the polarimetric Rayleigh scattering peak is accessible in most of the exo-Earths planetary systems.
%
The rainbow due to water clouds at phase angles of ${\sim}\qtyrange{20}{60}{\degree}$ would be accessible with \HWO\ for a planet with an Earth equivalent instellation in ${\sim}\num{46}$ systems, while the ocean glint signature at phase angles of ${\sim}\qtyrange{130}{170}{\degree}$ would be accessible in ${\sim}\num{16}$ systems, assuming an \IWA${=}\qty{62}{\mas}$ ($3\lambda/D$).
%
Improving the \IWA${=}\qty{41}{\mas}$ ($2\lambda/D$) increases accessibility to rainbows and glints by factors of approximately 2 and 3, respectively.
%
By observing these scattering features, \HWO could detect a surface ocean and water cycle, key indicators of habitability.
\looseness=-1
\end{abstract}

@@

In this work, we consider each orbit to have a semi-major axis for which it would receive the same flux as Earth assuming a circular orbit.
%
As Earth is at the inner edge of the Habitable Zone, our estimates for the number of systems with accessible scattering features is conservative.
%
However, as we are only considering the accessibility of scattering features, it is possible to scale the results presented here for different semi-major axis. 
%
This is because the range of accessible features is determined only by the amount of the orbit covered by the mask, therefore increasing the semi-major axis of an orbit (e.g., double) is equivalent to decreasing the apparent mask size by the same factor (e.g., half).
\looseness=-1

@@

The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
%
The instrument design and capabilities are still under development and should be driven by the mission's science goals.
%
In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarised and unpolarised reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability. 
%
We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
%
While we do not consider the detectability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light, but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.

>>>>

\begin{abstract}
NASA is engaged in planning for a Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones and establish their habitability. 
%
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
%
Direct imaging missions are optimised for planets near quadrature, but \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
%
The range of accessible phase angles for a given exoplanet will depend on the planet's orbital inclination and/or the coronagraph's inner working angle (IWA). 
%
We use a recently-created catalog relevant to HabWorlds of 164 stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and polarization effects due to Rayleigh scattering. 
%
We find that the polarimetric Rayleigh scattering peak is accessible in most of the exo-Earth planetary systems.
%
The rainbow due to water clouds at phase angles of ${\sim}\qtyrange{20}{60}{\degree}$ would be accessible with \HWO\ for a planet with an Earth equivalent instellation in ${\sim}\num{46}$ systems, while the ocean glint signature at phase angles of ${\sim}\qtyrange{130}{170}{\degree}$ would be accessible in ${\sim}\num{16}$ systems, assuming an \IWA${=}\qty{62}{\mas}$ ($3\lambda/D$).
%
Improving the \IWA${=}\qty{41}{\mas}$ ($2\lambda/D$) increases accessibility to rainbows and glints by factors of approximately 2 and 3, respectively.
%
By observing these scattering features, \HWO could detect a surface ocean and water cycle, key indicators of habitability.
\looseness=-1
\end{abstract}

@@

In this work, we consider each orbit to have a semi-major axis for which it would receive the same flux as Earth assuming a circular orbit.
%
As Earth is at the inner edge of the Habitable Zone, our estimates for the number of systems with accessible scattering features is conservative.
%
However, as we are only considering the accessibility of scattering features, it is possible to scale the results presented here for different semi-major axes. 
%
This is because the range of accessible features is determined only by the amount of the orbit covered by the mask, therefore increasing the semi-major axis of an orbit (e.g., double) is equivalent to decreasing the apparent mask size by the same factor (e.g., half).
\looseness=-1

@@

The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
%
The instrument design and capabilities are still under development and should be driven by the mission's science goals.
%
In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarised and unpolarised reflected light, such as rainbows, ocean glint, and glories, which could establish a planet's habitability. 
%
We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
%
While we do not consider the detectability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light, but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
